---
title: "Predicting Rent using Costs of Living - Multiple Linear Regression"
author: "Jennifer Mac"
date: "`r format(Sys.time(), '%B %d, %Y')`"
toc: true
execute: 
  warning: FALSE
  message: FALSE
format: 
  html:
    embed-resources: true
---

## Load Packages

```{r}
pacman::p_load(car, emmeans, tidyverse, knitr, leaps, janitor, corrplot, cowplot)
```

## Set Themes

```{r}
# DARK BLUE 
inside_plot_color <- "#e9ebf5ff"
outside_plot_color <- "#2c3151ff"
major_grid_color <- "white"
minor_grif_color <- "white"
text_color <- "white"
```

```{r}
# BLUE 
inside_plot_color <- "#e9ebf5ff"
outside_plot_color <- "#4b548bff"
major_grid_color <- "white"
minor_grif_color <- "white"
text_color <- "white"
```

```{r}
# white BLUE 
inside_plot_color <- "#cde2eaff"
outside_plot_color <- "white"
major_grid_color <- "white"
minor_grif_color <- "#e5f0f4ff"
text_color <- "#2c3151ff"
```

```{r}
# WHITE
inside_plot_color <- "#e9ebf5ff"
outside_plot_color <- "white"
major_grid_color <- "white"
minor_grif_color <- "white"
text_color <- "#2c3151ff"
```

```{r}
theme_color <- theme_minimal() +
  theme(panel.background = element_rect(fill = inside_plot_color, 
                                        colour = inside_plot_color),
        plot.background = element_rect(fill = outside_plot_color), 
        panel.grid.major = element_line(color = major_grid_color),
        panel.grid.minor = element_line(color = minor_grif_color),
        plot.title = element_text(face = "bold", 
                                  color = text_color, 
                                  family  = "mono"), 
        axis.title.x = element_text(face = "bold", 
                                    color = text_color, 
                                  family  = "mono"), 
        axis.title.y = element_text(face = "bold", 
                                    color = text_color, 
                                  family  = "mono"),
        axis.text.x = element_text(hjust = 0, 
                                   face = "bold", 
                                   margin = margin(l = -50),
                                   color = text_color, 
                                   family  = "mono", 
                                   size = 10), 
        axis.text.y = element_text(face = "bold", 
                                   color = text_color, 
                                   family  = "mono", 
                                   size = 10), 
        legend.position = "none")
```

## Countries

```{r}
countries <- c("Nigeria", "Hong Kong (China)", "Singapore", "Bangladesh", 
               "Barbados", "Kuwait", "Oman", "Saudi Arabia", "India", 
               "Cuba", "United States")
```

## Load Data (EDA)

+----------------------------------------------------+
| Variables                                          |
+----------------------------------------------------+
| Cost of living Index (Excluding Rent)              |
+----------------------------------------------------+
| Rent Index                                         |
+----------------------------------------------------+
| Local Purchasing Power Index                       |
|                                                    |
| (5,268.57/(3,533.4+1 **bedroom outside city**))/   |
|                                                    |
| (5,384.49/(6,076.2+**bedroom outside city**))\*100 |
+----------------------------------------------------+
| Food Index                                         |
|                                                    |
| (Grocery Index / Restaurant Index) \* 100          |
+----------------------------------------------------+
| Relative Purchasing Power Index                    |
|                                                    |
| (Local Purchasing Power / (cost of living + rent)) |
+----------------------------------------------------+

```{r}
# Cost of Living Indices with 121 countries
dat <- clean_names(read.csv("wikipedia_cost_of_living_indices3.csv"))
dat1 <- data.frame(dat, 
                   "food_index" = round(dat$groceries_index/dat$restaurant_price_index, 3)*100, # created due to colinearity
                   "relative_purchasing_power_index" = round(dat$local_purchasing_power_index/(dat$rent_index), 3))

# Cost of Living Indices with 404 countries
dat_city <- clean_names(read.csv("cost_of_living_cities.csv"))
dat_city1 <- data.frame(dat_city[, -c(4:6)], 
                        "food_index" = round(dat_city$groceries_index/dat_city$restaurant_price_index, 3)*100, 
                        "relative_purchasing_power_index" = round(dat_city$local_purchasing_power_index/dat_city$cost_of_living_plus_rent_index, 3)*100)
```

## Overview of Dataset - Looking for Outliers

```{r, fig.width=12, fig.height=6}
dat_clean <- dat1 |> select(-cost_of_living_plus_rent_index,
                            -restaurant_price_index, 
                            -groceries_index)
summary(dat_clean)

df_long <- tidyr::pivot_longer(dat_clean[, -c(1, 6)], everything())
df_long$name <- factor(df_long$name, levels = c("rent_index", "cost_of_living_index", 
                          "local_purchasing_power_index", "food_index"))

ggplot(df_long, aes(x = name, y = value)) + geom_boxplot(color = '#4b548bff') + 
  labs(title = "Summary Statistics", x = '') + theme_color + 
  theme(text = element_text(size = 18), 
        axis.text.x = element_text(face = "bold", color = text_color, family  = "mono"), 
        axis.text.y = element_text(face = "bold", color = text_color, family  = "mono")) +
  scale_x_discrete(labels = c('Rent','Cost of Living', 'Local Purchasing Power', 'Food'))

  
```

## Creating Linear Models

### Without Relative Purchasing power

log(`rent_index`) = -2.096 + 1.269 log(`cost_of_living`) + 0.003 (`local_purchasing_power_index`) - 0.002 (`food_index`)

```{r}

# full model: all variables as predictors 
mod_full <- lm(rent_index ~ ., dat1[, -c(1, 9)]); summary(mod_full)

corrplot(cor(dat1[, -c(1, 9)])*abs(cor(dat1[, -c(1, 9)]))>0.7) # colinearity
dat2 <- dat1[, -c(4:6)] # remove cost of living plus rent index, grocery index and restaurant index

corrplot(cor(dat2[, -c(1, 6)])*abs(cor(dat2[, -c(1, 6)]))>0.7) # no colinearity

# full model 2: remaining variables as predictors 
mod_full2 <- lm(rent_index ~ ., dat2[, -c(1, 6)]); summary(mod_full2)
par(mfrow = c(1, 2)); plot(mod_full2, 1:2)

#transform cost of living, local purchasing poer, and relative purchasing power
summary(powerTransform(cbind(cost_of_living_index, 
                             local_purchasing_power_index,
                             food_index)~1, dat1))

summary(powerTransform(cbind(cost_of_living_index, 
                             local_purchasing_power_index,
                             food_index)~1, dat1))

# AIC: backward and both are the same result
summary(step(mod_full2, direction = "forward"))

mod1 <- lm(log(rent_index) ~ log(cost_of_living_index) + 
             local_purchasing_power_index + 
             food_index, dat1)
par(mfrow = c(1, 2)); plot(mod1, 1:2); summary(mod1) # R^2 = 0.7722
```

```{r}
dat1$label_c <- ifelse(dat1$country %in% countries, dat1$country, '')
dat1$color_yn <- ifelse(dat1$country %in% countries, 'yes', 'no')
dat1$residuals <- residuals(mod1)
dat1$fitted_values <- fitted(mod1)

no_color <- "#9fa5c8ff"
yes_color <- "#2c3151ff"
inside_text_color <- "#2c3151ff"

fr <- ggplot(dat1, aes(x = fitted_values, y = residuals, label = label_c)) +
  geom_point(aes(color = color_yn)) +
  scale_color_manual(values = c("no" = no_color,
                                "yes" = yes_color)) +
  geom_text(nudge_y = -1, color = inside_text_color, fontface = 2, size = 2.5) +
  geom_hline(yintercept = 0, linetype = "dashed", scolor = inside_text_color) +
      labs(title = "Residuals vs Fitted Values",
           x = "Fitted Values",
           y = "Residuals") + theme_color 


qq <- ggplot(dat1, aes(sample = log(local_purchasing_power_index), label = label_c)) +
  stat_qq(geom = "point", color = ifelse(dat1$color_yn == "no", 
                                         no_color, inside_text_color)) +
  stat_qq_line(linetype = "dashed", color = inside_text_color) +
  labs(title = "Q-Q Residuals",
           x = "Standardized Residuals",
           y = "Theoretical Quantiles") + theme_color

p <- plot_grid(fr, qq);p
ggsave("plot.png", p, width = 7, height = 5)
```

```{r, fig.width=10, fig.height=8.2}
options(scipen = 999)

pred_rent_country <- data.frame(
  'country' = dat1$country,
  'pred' = predict(mod1), 
  'rent' = log(dat1$rent_index)
  ) |>
  mutate(error = round((pred/rent)-1, 6))
 
by_error <- pred_rent_country |> arrange(abs(error)) |> select(country, error) |>
  mutate(error = case_when(abs(error) < 0.00001 ~ error,
                           abs(error) > 0.00001 ~ round(error, 4), 
                           TRUE ~ error))

by_error |> slice(c(1:10, 112:121)) |>
  ggplot(aes(x = reorder(country, abs(error)), y = error)) + 
  geom_hline(yintercept = 0, color = '#4b548bff', lwd = 0) +
  geom_col(fill = "#4b548bff", color = "#4b548bff") +
  theme_color +
  geom_label(aes(label = error), hjust = -0.1, angle = 90, 
             text.color = "#4b548bff", border.color = "white", 
             size = 5) +
  labs(title = 'Top 10 Most and Least Expected Country Rents: Error %', 
       x = '', y = 'Error %') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, 
                                   face = "bold", 
                                    color = text_color, 
                                  family  = "mono"), 
        axis.text.y = element_text(face = "bold", 
                                    color = text_color, 
                                  family  = "mono"),
        text = element_text(size = 17)) +
  coord_cartesian(ylim = c(NA, 1.08)) +
  geom_vline(xintercept = 10.5, color = "#4b548bff")
```

```{r, fig.width=8, fig.height=6}
# 84 - rent 36 - cost of living + rent
d <- dat_clean |> select(country, relative_purchasing_power_index) |>
  arrange(desc(relative_purchasing_power_index)) |> slice(c(1:10, 84, 112:121)) 

plot_affordability <- d |>
  ggplot(aes(x = reorder(country, -relative_purchasing_power_index), 
             y = relative_purchasing_power_index, 
             fill = country == "United States")) + 
  geom_vline(xintercept = 11, color = "#4b548bff", lwd = 3) +
  geom_hline(yintercept = 0, color = '#4b548bff', lwd = 0) +
  geom_col(aes()) +
  scale_fill_manual(values = c("TRUE" = "#4b548bff", "FALSE" = "#2c3151ff"),
                    guide = "none") +
  theme_color +
  labs(title = 'Top 10 Most and Least Affordable Housing by Country:', 
       x = '', y = 'Relative Purchasing Power Index') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, 
                                   face = "bold", 
                                    color = text_color, 
                                  family  = "mono"), 
        axis.text.y = element_text(face = "bold", 
                                    color = text_color, 
                                  family  = "mono"),
        text = element_text(size = 15), 
        plot.margin = margin(t = 10, r = 60));plot_affordability

ggsave("plot_affordability.png", plot_affordability, width = 8.7, height = 5.5)
```

```{r, fig.width=12, fig.height=7}
no_color <- "#8bafbcff"
yes_color <- "#3b5a65ff"
inside_text_color <- "#3b5a65ff"
line_color <- "#3b5a65ff"

countries <- c("Nigeria", "Hong Kong (China)", "Singapore", "Bangladesh", 
               "Barbados", "Kuwait", "Oman", "Saudi Arabia", "India", 
               "Cuba", "Cameroon", "United States")

pred_rent_country <- data.frame(
  'country' = dat1$country,
  'pred' = predict(mod1), 
  'rent' = log(dat1$rent_index)
  ) |>
   mutate(error = round((pred/rent)-1, 6))

by_error <- pred_rent_country |> arrange(error) |> select(country, error, pred)
which(by_error$country %in% c("Hong Kong (China)", "Singapore","United States", "India", "Cuba", "Switzerland", "Nigeria", 
                              "Jordan", "Algeria", "Venezuela", "Pakistan", "Nepal", "Bangladesh", "Saudi Arabia", "Norway", "Denmark"))
countries <- by_error |> slice(1, 3, 5, 36, 61:62, 82, 91, 94, 109, 116:121)


pred_rent_country <- data.frame(
  'country' = dat1$country,
  'pred' = predict(mod1), 
  'rent' = log(dat1$rent_index)
  ) |>
  mutate(country = ifelse(country %in% countries$country, dat1$country, NA)) |>
  mutate(color_yn = ifelse(!is.na(country), 'yes', 'no'), 
           alpha_yn = ifelse(!is.na(country), 1, 0.7))

p <- pred_rent_country |>
  ggplot(aes(pred, rent, label = country)) + 
  geom_point(aes(color = color_yn), size = 3) +
  scale_color_manual(values = c("no" = no_color,
                                "yes" = yes_color))+
  geom_abline(intercept = 0, slope = 1, lwd = 1.5,
              linetype = "dashed", color = line_color)+
  ggrepel::geom_label_repel(color = line_color,
                            fontface = 2, family = "mono", size = 5, 
                            label.size = 0, box.padding = 0.1, 
                            min.segment.length = 0) +
  labs(x = "Predicted Rent", y = "Actual Rent", 
       title = "Predicted Rent vs Actual Rent: Countries") + 
  theme_color +
  theme(plot.margin = margin(t = 10, r = 50, b = 20, l = 20), 
        text = element_text(size = 20)); p


ggsave("p.png", p, width = 10, height = 6)
```

```{r, fig.width=10, fig.height=6.2}
mod2_city <- lm(log(rent_index) ~ log(cost_of_living_index) + 
           local_purchasing_power_index + food_index, dat_city1)

no_color <- "#9fa5c8ff"
yes_color <- "#2c3151ff"
inside_text_color <- "#2c3151ff"
line_color <- "#2c3151ff"

pred_rent <- data.frame('city' = dat_city1$city,
                        'pred' = predict(mod2_city), 
                        'rent' = log(dat_city1$rent_index)) |>
  mutate(error = round((pred/rent)-1, 6))

by_error <- pred_rent |> arrange(error) |> select(city, error, pred)
which(by_error$city %in% c("San Jose, CA, United States"))
city_names <- by_error |> slice(c(46, 41, 1:3, 402:404, 26, 46, 340, 351, 364, 365, 393, 398, 400, 8,  10, 188))

pred_rent <- data.frame('city' = dat_city1$city,
                        'pred' = predict(mod2_city), 
                        'rent' = log(dat_city1$rent_index)) |>
  mutate(city = ifelse(city %in% city_names$city, dat_city1$city, NA)) |>
  mutate(color_yn = ifelse(!is.na(city), 'yes', 'no'), 
         alpha_yn = ifelse(!is.na(city), 1, 0.7))

p <- pred_rent |>
  ggplot(aes(pred, rent, label = city)) + 
  geom_point(aes(color = color_yn), size = 3) +
  scale_color_manual(values = c("no" = no_color,
                                "yes" = yes_color))+
  geom_abline(intercept = 0, slope = 1, lwd = 1.5,
              linetype = "dashed", color = line_color)+
  ggrepel::geom_label_repel(color = line_color,
                            fontface = 2, family = "mono", size =3,  
                            label.size = 0, box.padding = 0.1, 
                            min.segment.length = 0) +
  labs(x = "Predicted Rent", y = "Actual Rent", 
       title = "Predicted Rent vs Actual Rent: Global Cities") + 
  theme_color +
  theme(plot.margin = margin(t = 10, r = 50, b = 20, l = 20), 
        text = element_text(size = 15)); p

ggsave("p.png", p, width = 8.7, height = 5.5)
```

```{r}

d <- dat_city1 |> select(city, relative_purchasing_power_index) |>
  arrange(desc(relative_purchasing_power_index))

which(d$city %in% c("San Jose, CA, United States"))

d <- d |> slice(c(1:5, 348, 400:404)) 
plot_affordability <- d |>
  ggplot(aes(x = reorder(city, -relative_purchasing_power_index), 
             y = relative_purchasing_power_index, 
             fill = city == "San Jose, CA, United States")) + 
  geom_vline(xintercept = 6, color = "#4b548bff", lwd = 3) +
  geom_hline(yintercept = 0, color = '#4b548bff', lwd = 0) +
  geom_col(aes()) +
  scale_fill_manual(values = c("TRUE" = "#4b548bff", "FALSE" = "#2c3151ff"),
                    guide = "none") +
  theme_color +
  labs(title = 'Top 5 Most and Least Affordable Housing by Global Cities:', 
       x = '', y = 'Relative Purchasing Power Index') +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, 
                                   face = "bold", 
                                    color = text_color, 
                                  family  = "mono"), 
        axis.text.y = element_text(face = "bold", 
                                    color = text_color, 
                                  family  = "mono"),
        text = element_text(size = 13), 
        plot.margin = margin(l= 30,t = 10, r = 30));plot_affordability

ggsave("plot_affordability.png", plot_affordability, width = 8.7, height = 5.3)
```

### SPARKLINES

```{r}
# SET UP
countries <- c("Nigeria", "Hong Kong (China)", "Singapore", "Bangladesh", 
               "Barbados", "Kuwait", "Oman", "Saudi Arabia", "India", 
               "Cuba", "Cameroon", "United States", "Max", "Min")
countries <- c("Nigeria", "Singapore", "Bangladesh", 
               "Barbados", "Kuwait", "India", 
               "Cuba", "United States", "Max", "Min")

min <- data.frame(country = "Min", 
                      cost_of_living_index = 18.8, 
                      rent_index = 2.4 ,
                      local_purchasing_power_index = 2.3,
                      food_index = 79.7)
max <- data.frame(country = "Max", 
                      cost_of_living_index = 101.1, 
                      rent_index = 67.2 ,
                      local_purchasing_power_index = 182.5,
                      food_index = 266.5)
new_dat <- rbind(rbind(dat1[, -c(4:6, 9:13)], max), min)[c(1:3, 5, 4)]
```

```{r}
new_dat <- new_dat |>
  mutate(
    Country = case_when(
    country == "Nigeria" ~ "Nigeria",
    country == "Hong Kong (China)" ~ "Hong Kong",
    country == "Singapore" ~ "Singapore",
    country == "Bangladesh" ~ "Bangladesh",
    country == "Barbados" ~ "Barbados",
    country == "Kuwait" ~ "Kuwait",
    country == "Oman" ~ "Oman",
    country == "Saudi Arabia" ~ "Saudi Arabia",
    country == "India" ~ "India",
    country == "Cuba" ~ "Cuba",
    country == "Cameroon" ~ "Cameroon",
    country == "United States" ~ "United States",
    TRUE  ~ '.'),
    alphaLevel = case_when(
      Country != '.' ~ 1, 
      TRUE ~ 0.1))

new_dat <- new_dat |>
  mutate(
    Country = case_when(
    country == "Nigeria" ~ "Nigeria",
    country == "Singapore" ~ "Singapore",
    country == "Bangladesh" ~ "Bangladesh",
    country == "Barbados" ~ "Barbados",
    country == "Kuwait" ~ "Kuwait",
    country == "India" ~ "India",
    country == "Cuba" ~ "Cuba",
    country == "United States" ~ "United States",
    TRUE  ~ '.'),
    alphaLevel = case_when(
      Country != '.' ~ 1, 
      TRUE ~ 0.1))
```

```{r}
label_food <- new_dat |>
  mutate(side = "food_index", 
         scaled_food = (food_index - 
                         min(food_index)) /
                 (max(food_index) - 
                    min(food_index))) |>
  filter(Country != '.') |>
  select(Country, side, scaled_food) 

new_dat <- new_dat |> filter(country %in% countries) 
new_dat <- new_dat |> select(country, rent_index, cost_of_living_index, local_purchasing_power_index,
                         food_index, Country, alphaLevel)

color <- data.frame(color = c("#2c3151ff","#4b548bff", "#769faeff",
                              "#cfc79dff", "#D0D4E6", "red4"))
```

```{r, fig.width=8, fig.height=4.5}
# PLOT 
new_dat1 <- new_dat %>%
  arrange(Country %in% c("Bangladesh", "Singapore", "Nigeria", "United States"))

my_plot <- new_dat1 |>
  GGally::ggparcoord(columns = 2:5, 
                     groupColumn = 6, 
                     scale = "uniminmax", 
                     alphaLines = 'alphaLevel',
                     title = "INDEXES:") +
  scale_color_manual(values = c("Bangladesh" = color[5, 1], 
                                "Barbados" = color[5, 1],
                                "United States" = color[6, 1],
                                "India" = color[1, 1],
                                "Nigeria" = color[5, 1], 
                                "Oman" = color[5, 1], #paired2
                                "Singapore" = color[5, 1], #paired1
                                "Hong Kong" = color[5, 1], #paired1
                                "Saudi Arabia" = color[5, 1], #paired2
                                "Kuwait" = color[5, 1],
                                "Cuba" = color[5, 1], #paired3
                                "Cameroon" = color[5, 1], #paired3
                                "." = NA)) +
  geom_line(size = 1.5, lineend = "round", 
          aes(linetype = Country, linewidth = Country)) +
  scale_linetype_manual(values = c(
      "United States" = "dashed",
      "Bangladesh"    = "solid",
      "Nigeria"       = "solid",
      "Singapore"     = "solid",
      "Hong Kong"     = "solid", 
      "Barbados"      = "solid",
      "India"         = "solid",
      "Oman"          = "solid",
      "Saudi Arabia"  = "solid",
      "Kuwait"        = "solid",
      "Cuba"          = "solid",
      "Cameroon"      = "solid"
    )) +
  scale_linewidth_manual(values = c(
      "Bangladesh"    = 1, 
      "Barbados"      = 1,
      "United States" = 1,
      "India"         = 1,
      "Nigeria"       = 1,
      "Oman"          = 1,
      "Singapore"     = 1,
      "Hong Kong"     = 1,
      "Saudi Arabia"  = 1,
      "Kuwait"        = 1,
      "Cuba"          = 1,
      "Cameroon"      = 1)) +
  labs(x = "Variables", y = "Scaled Value (0-1)") +
  guides(alpha = "none") +
  scale_x_discrete(labels = c('•Rent', '•Cost of Living', '•Purchasing Power', '•Food'), 
                   expand = expansion(add = c(0.02, 0.8))) + 
  theme_color +
  theme(plot.margin = margin(t = 10, r = 20, b = 20, l = 15)) + 
  geom_point(size = 2)

my_plot1 <- my_plot +
  ggrepel::geom_label_repel(data = label_food,
                            aes(x = "food_index", 
                                y = scaled_food,
                                label = Country,
                                color = Country), 
                            inherit.aes = FALSE,
                            hjust = "left", direction = "y", min.segment.length = 0, 
                            size = 4, label.size = 0, nudge_x = 0.5, box.padding = 0.2, 
                            fontface = "bold", family = "mono", xlim = 4, , seed = 27); my_plot1

ggsave("my_plot.png", my_plot1, width = 8, height = 4.5)
```

WITH RELATIVE PURCHASING POWER

```{r}
# forward step with AIC (all step models are the same)
mod3 <- lm(log(rent_index) ~ log(cost_of_living_index) + 
             local_purchasing_power_index + 
             relative_purchasing_power_index + 
             food_index, dat1[, -1]); summary(mod3)

# no colinearity
corrplot(cor(dat1[, -c(1, 10:13)])*abs(cor(dat1[, -c(1, 10:13)]))>0.7)
par(mfrow = c(1, 2)); plot(mod3, 1:2)

# AIC: backward and both are the same result
summary(step(mod3, direction = "forward"))

#transform cost of living, local purchasing poer, and relative purchasing power
summary(powerTransform(cbind(cost_of_living_index, 
                             local_purchasing_power_index,
                             relative_purchasing_power_index, 
                             food_index)~1, dat1))

mod4 <- lm(log(rent_index) ~ log(cost_of_living_index) + 
             local_purchasing_power_index + 
             log(relative_purchasing_power_index) + 
             food_index, dat1[, -1]); summary(mod4)

par(mfrow = c(1, 2)); plot(mod4, 1:2)
```

GRAPH Median income to median House sale price US

```{r, fig.width=8, fig.height=4.5}
income_housing <- left_join(read.csv("MEHOINUSA672N.csv"), read.csv("MSPUS.csv"), by = "observation_date") |>
  rename(house_price = MEHOINUSA672N) |> rename(income = MSPUS) |>
  mutate(observation_date = as.numeric(substr(observation_date, 1, 4))) |>
  filter(observation_date >= 1985)|> rename(year = observation_date)
```

```{r, fig.width=10, fig.height=6.2}
# Cost of Living Indices with 121 countries
dat <- clean_names(read.csv("US_cities.csv"))
dat1 <- dat |> mutate(city = str_remove(city, ", United States"), 
                      food_index = round(groceries_index / restaurant_price_index, 3) * 100,
                      relative_purchasing_power_index = round(local_purchasing_power_index / rent_index, 3))

dat1 <- dat |> mutate(city = str_remove(city, ", United States"), 
                      food_index = round(groceries_index / restaurant_price_index, 3) * 100,
                      relative_purchasing_power_index = round(local_purchasing_power_index / (rent_index+cost_of_living_index), 3))

mod1 <- lm(log(rent_index) ~ log(cost_of_living_index) + 
             local_purchasing_power_index + 
             food_index, dat1)

pred_rent_country <- data.frame(
  'city' = dat1$city,
  'pred' = predict(mod1), 
  'rent' = log(dat1$rent_index)
  ) |>
   mutate(error = round((pred/rent)-1, 6))

by_error <- pred_rent_country |> arrange(error) |> select(city, error, pred) 
which(by_error$city %in% c("New York, NY", "San Francisco, CA", "San Jose, CA", "Honolulu, HI"))

norm_ab_UScity <- by_error |> slice(c(1:2, 70, 73:75, 14, 22, 31, 69))

pred_rent_country <- data.frame(
  'city' = dat1$city,
  'pred' = predict(mod1), 
  'rent' = log(dat1$rent_index)
  ) |>
  mutate(city = ifelse(city %in% norm_ab_UScity$city, 
                          dat1$city, NA)) |>
  mutate(color_yn = ifelse(!is.na(city), 'yes', 'no'), 
           alpha_yn = ifelse(!is.na(city), 1, 0.7))

no_color <- "#9fa5c8ff"
yes_color <- "#2c3151ff"
inside_text_color <- "#2c3151ff"
line_color <- "#2c3151ff"

US_city_plot <- pred_rent_country |>
  ggplot(aes(pred, rent, label = city)) + 
  geom_point(aes(color = color_yn), size = 3) +
  scale_color_manual(values = c("no" = no_color,
                                "yes" = yes_color))+
  geom_abline(intercept = 0, slope = 1, lwd = 1.5,
              linetype = "dashed", color = line_color)+
  ggrepel::geom_label_repel(color = line_color,
                            fontface = 2, family = "mono", size = 4, 
                            label.size = 0, box.padding = 0.3, 
                            min.segment.length = 0) +
  labs(x = "Predicted Rent", y = "Actual Rent", 
       title = "Predicted Rent vs Actual Rent: US Cities") + 
  theme_color +
  theme(plot.margin = margin(t = 10, r = 50, b = 20, l = 20), 
        text = element_text(size = 15)); US_city_plot

ggsave("US_city_plot.png", US_city_plot, width = 8.7, height = 5.5)
```

```{r}
dat_clean <-  dat1 |> select(-cost_of_living_plus_rent_index,
                            -restaurant_price_index, 
                            -groceries_index)


d <- dat_clean |> select(city, relative_purchasing_power_index) |>
  arrange(desc(relative_purchasing_power_index)) |> slice(c(1:5, 70:75)) 

plot_affordability <- d |>
  ggplot(aes(x = reorder(city, -relative_purchasing_power_index), 
             y = relative_purchasing_power_index, 
             fill = city == "San Jose, CA")) + 
  geom_vline(xintercept = 6, color = "#4b548bff", lwd = 3) +
  geom_hline(yintercept = 0, color = '#4b548bff', lwd = 0) +
  geom_col(aes()) +
  scale_fill_manual(values = c("TRUE" = "#4b548bff", "FALSE" = "#2c3151ff"),
                    guide = "none") +
  theme_color +
  labs(title = 'Top 5 Most and Least Affordable Housing by US Cities:', 
       x = '', y = 'Relative Purchasing Power Index') +
  theme(text = element_text(size = 15), 
        plot.margin = margin(t = 10, r = 60));plot_affordability

ggsave("plot_affordability.png", plot_affordability, width = 8.7, height = 5.5)
```
